---
title: "Education and Distance."
author: "gregory_ut"
date: '2022-01-23'
slug: []
categories:
- R
- sf
- ggplot2
- dplyr
tags:
- ggplot2
- map
- R Markdown
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show = "static")
library(tidyverse)
library(here)
library(tidycensus)
library(sf)
```

# This blog will explore the relationship between distance from Central Business District and proportion of different groups by educational attainment status.

### Thank you by Kyle E. Walker for providing guiding code instructions. [link](https://link.springer.com/article/10.1007/s40980-018-0042-7)


## First get latest American Community Survey Data.
```{r}

ny_df <-
  get_acs(
    geography = "tract",
    variables = c(hs = "B15003_017", grad = "B15003_023"),
    state = "NY",
    summary_var = "B15003_001",
    geometry = TRUE
  )
```

## Get spatial file of core based statistical areas
```{r, results='hide'}
# get CBSA shapefile
cbsa <- tigris::core_based_statistical_areas(cb = TRUE) %>%
  select(GEOID, metro_name = NAME)
```

## Spatial join ACS data with CBSA file
```{r}

ny_sp <- ny_df %>%
  st_join(cbsa, join = st_within) %>%
  filter(
    str_detect(metro_name, "New York-Newark-Jersey City"),!str_detect(NAME, "Suffolk|Rockland|Westchester|Putnam|Pike")
  )
```

## Create Location Quotient function and city hall coordinates

$$LQ = \frac{G_i / T_i}{G_m / T_m}$$
Where $i$ is a census tract, $m$ is the metro area, $G$ is group population and $T$ is the total population.

```{r}
loc_quot <- function(group, total) {
  num <- group / total
  denom <- sum(group, na.rm = TRUE) / sum(total, na.rm = TRUE)
  output <- num / denom
  return(output)
}

city_hall <- c(-74.0061, 40.7128) %>%
  st_point() %>%
  st_sfc(crs = st_crs(ny_sp)) %>%
  st_transform(4269)

```

## Take location quotient for every group and then create distance between each tract to city hall.
```{r}
ny_df <- ny_sp %>%
  group_by(variable) %>%
  mutate(lq = loc_quot(estimate, summary_est)) %>%
  ungroup() %>%
  st_transform(4269) %>%
  mutate(dist = as.numeric(st_distance(st_centroid(.), city_hall)))

ny_df %>%
  glimpse()

```

## Finally plot!
```{r, warning=FALSE}


ny_df %>%
  ggplot(aes(dist / 1000, lq, color = variable)) +
  geom_smooth(method = 'loess', span = 0.3) +
  theme_minimal() +
  labs(
    color = "Degree",
    x = "Distance (km)",
    y = "LQ",
    title = "As Distance From City Hall Increases,\n % With Degrees Fluctuates."
  ) +
  scale_color_manual(
    values = c('#1D9EF5', '#3FF894'),
    labels = c("Graduate Degree", "High School Degree")
  ) +
  geom_hline(
    yintercept = 1,
    color = 'black',
    linetype = 'dashed',
    alpha = 0.5
  ) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 17, face = 'bold', hjust = 0.5),
    legend.position = 'bottom'
  )

```

# Now some PURRR, for multiple cities, I am doing 2 but any number is possible, as long as the projection is right.
```{r, results = "hide"}


state <- c("New York", "Pennsylvania")

acs_df <-
  state %>%
  map_df(., function(x) {
    get_acs(
      geography = "tract",
      variables = c(hs = "B15003_017", grad = "B15003_023"),
      state = x,
      summary_var = "B15003_001",
      geometry = TRUE
    )
  })

glimpse(acs_df)
```

## Filter out unwated geographies and grab state names. 
```{r}
ny_ph_cbsa <- acs_df %>%
  st_join(cbsa, join = st_within) %>%
  filter(
    str_detect(metro_name, "New York-Newark-Jersey City|Philadelphia"),!str_detect(NAME, "Suffolk|Rockland|Westchester|Putnam|Pike")
  ) %>%
  mutate(
    county = str_extract(NAME, regex("(\\w+) county, (\\w.*)", ignore_case = TRUE)),
    state = str_extract(county, regex(", (\\w.*)")),
    state = str_remove(state, ", ")
  )


ny_ph_cbsa %>%
  ggplot() +
  geom_sf(lwd = 0.1) +
  theme_void()

```

## Get city hall coordinates for both cities and make it spatial. 
```{r, warning=FALSE}
loc_quot <- function(group, total) {
  num <- group / total
  denom <- sum(group, na.rm = TRUE) / sum(total, na.rm = TRUE)
  output <- num / denom
  return(output)
}


city_halls <-
  tibble(
    long = c(-74.0061,-75.1635),
    lat = c(40.7128, 39.9528),
    state = c("New York", "Pennsylvania")
  )

cent_business <- city_halls %>%
  st_as_sf(coords = c("long", "lat"), crs = 4269)

# verify that city hall locations match
ny_ph_cbsa %>%
  ggplot() +
  geom_sf(lwd = 0.1) +
  geom_sf(data = cent_business, color = "red") +
  
  theme_void()

```

## Loop through each city and apply location quotient function.
```{r}


states <- c("New York", "Pennsylvania")

ny_ph_dist <- map_df(states, function(x) {
  metro <- ny_ph_cbsa %>%
    filter(state == x) %>%
    st_transform(4269)
  
  hall <- cent_business %>% filter(state == x)
  dist <- st_distance(st_centroid(metro), hall)
  
  mindist <- map(dist,  ~ min(.)) %>% unlist()
  metro_dist <- metro %>%
    mutate(dist_min = mindist)
  return(metro_dist)
})

```


## Plot !
```{r}
plot_df <- ny_ph_dist %>%
  group_by(state, variable) %>%
  mutate(lq = loc_quot(estimate, summary_est)) %>%
  ungroup()


plot_df %>%
  ggplot(aes(dist_min / 1000, lq, color = variable)) +
  geom_smooth(methods = "loess", span = 0.3) +
  facet_wrap( ~ state, scales = "free") +
  theme_minimal() +
  labs(
    color = "Degree",
    x = "Distance (km)",
    y = "LQ",
    title = "As Distance From City Hall Increases,\n % With Degrees Fluctuates."
  ) +
  scale_color_manual(
    values = c('#1D9EF5', '#3FF894'),
    labels = c("Graduate Degree", "High School Degree")
  ) +
  geom_hline(
    yintercept = 1,
    color = 'black',
    linetype = 'dashed',
    alpha = 0.5
  ) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 17, face = 'bold', hjust = 0.5),
    legend.position = 'bottom'
  )

```

