---
title: 'SQLite Github Challenge '
author: gutkin
date: '2022-12-22'
slug: []
categories:
  - Blog
tags:
  - sqlite3
  - blogdown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(tidyverse)
# 
# read_csv("https://raw.githubusercontent.com/AlexanderConnelly/BrainTree_SQL_Coding_Challenge_Data_Analyst/master/data_csv/continent_map.csv") %>%
#   write_csv(., "continent_map.csv")
# 
# read_csv("https://raw.githubusercontent.com/AlexanderConnelly/BrainTree_SQL_Coding_Challenge_Data_Analyst/master/data_csv/continents.csv") %>%
#   write_csv(., "continents.csv")
# 
# read_csv("https://raw.githubusercontent.com/AlexanderConnelly/BrainTree_SQL_Coding_Challenge_Data_Analyst/master/data_csv/countries.csv") %>%
#   write_csv(., "countires.csv")
# 
# read_csv("https://raw.githubusercontent.com/AlexanderConnelly/BrainTree_SQL_Coding_Challenge_Data_Analyst/master/data_csv/per_capita.csv") %>%
#   write_csv(., "per_capita.csv")

library(RSQLite)
library(DBI)
```

### All material came be found here: 
[link](https://github.com/AlexanderConnelly/BrainTree_SQL_Coding_Challenge_Data_Analyst)


First establish connection with the database.
If the table is too long i will LIMIT it by 20 to make things fit.
```{r}

con <- dbConnect(RSQLite::SQLite(), "C:\\Users\\grish\\Desktop\\sql-challenge\\my_db.db")

dbListTables(con)

dbGetQuery(con, "SELECT * FROM continents;")
dbGetQuery(con, "SELECT * FROM countries LIMIT 10;")
dbGetQuery(con, "SELECT * FROM continent_map LIMIT 10;")
dbGetQuery(con, "SELECT * FROM per_capita LIMIT 10;")


```

# question 1
1. Data Integrity Checking & Cleanup

Alphabetically list all of the country codes in the continent_map table that appear more than once. Display any values where country_code is null as country_code = "FOO" and make this row appear first in the list, even though it should alphabetically sort to the middle. Provide the results of this query as your answer.

For all countries that have multiple rows in the continent_map table, delete all multiple records leaving only the 1 record per country. The record that you keep should be the first one when sorted by the continent_code alphabetically ascending. Provide the query/ies and explanation of step(s) that you follow to delete these records.

# 1.1
```{r}


dbGetQuery(
  con,
  "WITH foo_tbl AS (SELECT continent_code,
           CASE WHEN country_code = 'NA' THEN 'FOO' ELSE country_code END AS country_code
  FROM continent_map)
  SELECT country_code,
  COUNT(*) as my_count
  FROM foo_tbl
  WHERE country_code = 'FOO'
  GROUP BY country_code
  HAVING COUNT(*) > 1
  UNION
  SELECT country_code,
           COUNT(*) AS my_count
           FROM continent_map
           WHERE country_code <> 'NA'
           GROUP BY country_code
           HAVING COUNT(*) > 1
  ORDER BY my_count DESC;")
```

# 1.2
```{r}
dbGetQuery(
  con,
  "WITH cte AS(
          SELECT cm.*,
          CASE WHEN country_code IS NULL THEN 'FOO' END AS country_code,
          ROW_NUMBER() OVER(PARTITION BY country_code ORDER BY country_code) AS dup_count 
  FROM continent_map AS cm)
          SELECT DISTINCT country_code, continent_code
          FROM cte WHERE dup_count == 1
          ORDER BY country_code
  LIMIT 20;"
)

```

2. List the countries ranked 10-12 in each continent by the percent of year-over-year growth descending from 2011 to 2012.

The percent of growth should be calculated as: ((2012 gdp - 2011 gdp) / 2011 gdp)

The list should include the columns:

rank
continent_name
country_code
country_name
growth_percent

# Part 2.1
```{r}

dbGetQuery(
  con,
  "WITH ranked_tbl AS(
           SELECT *,
           ctr.country_name,
           CASE WHEN pc.year = 2011 THEN gdp_per_capita ELSE NULL END AS gdp_2011,
           CASE WHEN pc.year = 2012 THEN gdp_per_capita ELSE NULL END AS gdp_2012
           FROM per_capita AS pc
           LEFT JOIN continent_map cm ON pc.country_code = cm.country_code
           LEFT JOIN continents cc ON cm.continent_code = cc.continent_code
           LEFT JOIN countries ctr ON pc.country_code = ctr.country_code
           WHERE pc.year IN(2011, 2012)),
          wide_tbl AS(
           SELECT continent_name,
            country_name,
           gdp_2011,
           gdp_2012,
           country_code
           FROM ranked_tbl
           ORDER BY country_code),
           deduped_tbl AS (
            SELECT
            wbl.continent_name,
            wbl.country_code,
            wbl.country_name,
             ((wb.gdp_2012 - wbl.gdp_2011) / wbl.gdp_2011) AS growth_percent
          FROM wide_tbl wbl
          LEFT JOIN wide_tbl wb ON wbl.country_code = wb.country_code
          WHERE wbl.gdp_2011 IS NOT NULL
          AND wb.gdp_2012 IS NOT NULL 
          AND growth_percent IS NOT NULL),
        my_tbl AS(
          SELECT
          DENSE_RANK() OVER(PARTITION BY continent_name ORDER BY growth_percent DESC) AS my_rank,
          continent_name,
          country_code,
          country_name, 
          growth_percent
  FROM deduped_tbl)
  SELECT my_rank, continent_name, country_code, country_name, growth_percent
  FROM my_tbl
  WHERE continent_name IS NOT NULL AND
  my_rank BETWEEN 10 AND 12;"
)
```


3. For the year 2012, create a 3 column, 1 row report showing the percent share of gdp_per_capita for the following regions:

(i) Asia, (ii) Europe, (iii) the Rest of the World. Your result should look something like...

# Part 3.1
```{r}

dbGetQuery(con, "with gdp_df as (SELECT *,
           CAST(gdp_per_capita AS int)
           FROM per_capita
           LEFT JOIN continent_map ON per_capita.country_code = continent_map.country_code
           LEFT JOIN continents ON continent_map.continent_code = continents.continent_code
           WHERE year = 2012),
           cc_df AS (
           SELECT continent_name, COUNT(*) as cc
           FROM gdp_df
           GROUP BY continent_name
           HAVING continent_name IS NOT NULL),
           cc2 AS(SELECT *, 
           CASE WHEN continent_name IN('Africa', 'North America', 'Oceania', 'South America') THEN 'Rest of World' ELSE continent_name END AS continent_name2
           FROM cc_df
           ),
           cc2_grp AS(
           SELECT continent_name2,
           SUM(cc) AS cc
           FROM cc2
           GROUP BY continent_name2),
           cc3_grp AS (SELECT 
           continent_name2,
           cc,
           SUM(cc) OVER() as pp
           FROM cc2_grp
           GROUP BY continent_name2)
           SELECT continent_name2,
           ROUND((cc / (pp * 1.00)) * 100, 0) as perc
           FROM cc3_grp;")


```

4a. What is the count of countries and sum of their related gdp_per_capita values for the year 2007 where the string 'an' (case insensitive) appears anywhere in the country name?
```{r}
dbGetQuery(con, "with gdp_df as (SELECT *
           FROM per_capita
           LEFT JOIN continent_map ON per_capita.country_code = continent_map.country_code
           LEFT JOIN continents ON continent_map.continent_code = continents.continent_code
           LEFT JOIN  countries ON per_capita.country_code = countries.country_code
           WHERE year = 2007)
           SELECT
           COUNT(*) OVER() as count,
           SUM(gdp_per_capita) OVER() as sum 
           FROM gdp_df
           WHERE country_name LIKE '%an%'
           LIMIT 1")

```

4b. Repeat question 4a, but this time make the query case sensitive.
```{r}
dbGetQuery(con, "with gdp_df as (SELECT *
           FROM per_capita
           LEFT JOIN continent_map ON per_capita.country_code = continent_map.country_code
           LEFT JOIN continents ON continent_map.continent_code = continents.continent_code
           LEFT JOIN  countries ON per_capita.country_code = countries.country_code
           WHERE year = 2007)
           SELECT
           COUNT(*) OVER() as count,
           SUM(gdp_per_capita) OVER() as sum 
           FROM gdp_df
           WHERE country_name LIKE '%an'
           LIMIT 1")
```


5. Find the sum of gpd_per_capita by year and the count of countries for each year that have non-null gdp_per_capita where (i) the year is before 2012 and (ii) the country has a null gdp_per_capita in 2012. Your result should have the columns:

year
country_count
total
```{r}
dbGetQuery(con, "SELECT year,
           COUNT(*) OVER(PARTITION BY country_name) as country_count,
           SUM(gdp_per_capita) AS total
           FROM per_capita
           LEFT JOIN continent_map ON per_capita.country_code = continent_map.country_code
           LEFT JOIN continents ON continent_map.continent_code = continents.continent_code
           LEFT JOIN countries ON per_capita.country_code = countries.country_code
           GROUP BY year
           HAVING gdp_per_capita NOT NULL AND
           year < 2012;")

```

6. All in a single query, execute all of the steps below and provide the results as your final answer:

a. create a single list of all per_capita records for year 2009 that includes columns:

continent_name
country_code
country_name
gdp_per_capita

b. order this list by:

continent_name ascending
characters 2 through 4 (inclusive) of the country_name descending

c. create a running total of gdp_per_capita by continent_name

d. return only the first record from the ordered list for which each continent's running total of gdp_per_capita meets or exceeds $70,000.00 with the following columns:

continent_name
country_code
country_name
gdp_per_capita
running_total
```{r}
dbGetQuery(con, "
                 WITH cdf AS (SELECT 
           continents.continent_name,
           continents.continent_code,
           countries.country_name,
           per_capita.gdp_per_capita,
           SUM(gdp_per_capita) OVER(PARTITION BY continent_name) AS running_total,
           ROW_NUMBER() OVER(PARTITION BY country_name ORDER BY gdp_per_capita) as row
           FROM per_capita
           LEFT JOIN continent_map ON per_capita.country_code = continent_map.country_code
           LEFT JOIN continents ON continent_map.continent_code = continents.continent_code
           LEFT JOIN countries ON per_capita.country_code = countries.country_code 
           WHERE continent_name IS NOT NULL
           ORDER BY continent_name,
           SUBSTRING(country_name, 2, 4) DESC)
           SELECT continent_name,
           country_code,
           countries.country_name,
           gdp_per_capita,
           running_total
           FROM cdf
           LEFT JOIN countries ON cdf.country_name = countries.country_name
           WHERE running_total >= 70000.00 AND
           row = 1
           LIMIT 100;")



```

7. Find the country with the highest average gdp_per_capita for each continent for all years. Now compare your list to the following data set. Please describe any and all mistakes that you can find with the data set below. Include any code that you use to help detect these mistakes.
```{r}
dbGetQuery(con, "WITH avg_df AS (SELECT 
           per_capita.year,
           continents.continent_code,
           continents.continent_name,
           countries.country_code,
           countries.country_name,
           AVG(per_capita.gdp_per_capita) OVER(PARTITION BY country_name) as avg_gdp
           FROM per_capita
           LEFT JOIN continent_map ON per_capita.country_code = continent_map.country_code
           LEFT JOIN continents ON continent_map.continent_code = continents.continent_code
           LEFT JOIN countries ON per_capita.country_code = countries.country_code
           WHERE continents.continent_name IS NOT NULL AND
           per_capita.gdp_per_capita IS NOT NULL),
           rank_df AS(SELECT *,
           RANK() OVER(PARTITION BY continent_name ORDER BY avg_gdp DESC) as my_rank
           FROM avg_df)
           SELECT my_rank,
           continent_name,
           country_code,
           country_name,
           avg_gdp
           FROM rank_df
           GROUP BY my_rank,
           continent_name,
           country_code,
           country_name,
           avg_gdp
           HAVING my_rank = 1
           LIMIT 20")

```

